---
title: "AqualogCompilation"
output: html_document
---

Sydney Shelton and Wes Slaughter 10/02/2023
This code is for compiling various different things from the Aqualog Data:
a254 for SUVA calculation, Spectral Slopes and Slope Ratio
BIX, HIX, FI, P/H, Protein-like, Humic-like to match the old fluorometer code with dilution correction
Inner filter effect correction, raman normalization, remove raman/rayleigh scattering lines, and calculate indices using staRdom

Note for setting up on fluoro computer - R markdown automatically sends each chunk to script location folder. You can change this to current 
working directory by going to Tools -> Global Options -> R Markdown -> Evaluate chunks in directory to "current"

Set the working directory to the folder for your run! Double check that all file names are correct and match, and then run the script.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list=ls()) 

# call Libraries
library(tidyverse)
library(staRdom)

## # Box connection stuff (ignore)
## library(boxr)
## library(rjson)

## # connect to Box
## config <- fromJSON(file = './config.json')

## box_auth(
##   client_id = config$client_id,
##   client_secret = config$client_secret)

```

#Organize all of the files into their respective folders
Create a folder for the high resolution absorbance data (which we use), the EEMs, and the low resolution absorbance data (for IFE correction)

```{r fileOrg}
# connect to raw data filepath
#This pull out the working directory, can add ", "data"" and it specifies a folder within the WD
fp.data <- file.path("C:/Users/shelts3/Dropbox/GradSchool/Research/MethodologiesGuidesAnalysisCode/lab-aqualog/20231030_ExampleRun_SASRoutineMonitoring")

# the final run results will be saved in:
fp.results <- file.path(fp.data, "results")
if(!dir.exists(fp.results)) {
  dir.create(fp.results) 
}

#List all of the files we'll read in
filenames <- list.files(fp.data, pattern="*.dat")

# seperate all of the absorbance data into "Absorbance" folder, all interpolated EEMS data into an "EEMS_IN" folder, 
# all non-interpolated EEMS into "EEMS_RAW", and an all data folder "all"
fp.abs <- file.path(fp.data, "Absorbance_highres")
if(!dir.exists(fp.abs)) {
  dir.create(fp.abs) 
}

#folder for interpolated EEMS
fp.eems.in <- file.path(fp.data, "EEMS_IN")
if(!dir.exists(fp.eems.in)) {
  dir.create(fp.eems.in) 
}

#folder for not interpolated eems
fp.eems.raw <- file.path(fp.data, "EEMS_RAW")
if(!dir.exists(fp.eems.raw)) {
  dir.create(fp.eems.raw) 
}

# folder for low res abs from eems
fp.lowa <- file.path(fp.data, "Absorbance_lowres")
if(!dir.exists(fp.lowa)) {
  dir.create(fp.lowa) 
}

# folder for combined abs and eems data
fp.all <- file.path(fp.data, "all")
if(!dir.exists(fp.all)) {
  dir.create(fp.all) 
}

for(fn in filenames) {
  # file path of raw file
  fp <- file.path(fp.data, fn)

  # set destination file path for abs and EEMS, and all
  fda <- file.path(fp.abs, fn)
  fde.in <- file.path(fp.eems.in, fn)
  fde.raw <- file.path(fp.eems.raw, fn)
  fdlowa <- file.path(fp.lowa, fn)
  fdall <- file.path(fp.all, fn)
  
  # copy absorbance data into absorbance folder
  if(grepl("- Abs Spectra Graphs", fn)) {
    print(paste("sorting data for: ", fn))

    # seperate high and low res
    if(grepl("^A", fn)) {
      file.copy(fp, fda)
    } else if(grepl("^E", fn)) {
      file.copy(fp, fdlowa)
    } else {
      ## print(paste("WARNING: no E or A prefix on ", fn, "cannot be sorted"))
    }
    
    file.copy(fp, fdall)
  }

  # copy EEMS_IN data into EEMS_IN folder
  if(grepl("Processed Graph_ IN", fn)) {
    file.copy(fp, fde.in)
    file.copy(fp, fdall)
  }
  
  # copy EEMS_RAW data into EEMS_RAW folder
  if(grepl("Sample - Blank Waterfall Plot", fn)) {
    file.copy(fp, fde.raw)
    file.copy(fp, fdall)
  }
}

```

#Absorbance formatting/calculations
Put all of the Napierian absorbance coefficients into one dataframe
Include Spectral slopes from 275-295 nm and 350-400 nm, and slope ratio (275–295-nm slope : 350–400-nm slope)
Plus the decadic form of a254 for SUVA calculation


```{r Absorbance}

#Read in all of the absorbance spectra ####
filenames <- list.files(fp.abs, pattern="*.dat", full.names = TRUE)

# then call them in
rawabs <- sapply(filenames, FUN = read.delim, simplify = FALSE, USE.NAMES = TRUE)

# Need to clean up the names in the list
# Change filenames to remove the canned part we don't want
# then make these object names in rawabs
filenames <- sub(" .*", "", basename(filenames))

names(rawabs) <- filenames

# Now, compile all of these columns into one dataframe ####
# Remove all columns except for the absorbance
rawabs <- lapply(rawabs, function(x) {select(x, -c("I", "I.dark", "R", "R.dark", "XCorrect", "Ic", "Rc", "Ic.Rc", "Percent.T"))})

#Make the Absorbance column name the sample name and remove extra columns
for(i in unique(1:length(filenames))){
  colnames <- c("Wavelength", filenames[i])
  colnames(rawabs[[i]]) <- colnames
}

#Combine into one dataframe, joining by Wavelength column
spectra <- reduce(rawabs, left_join, by = "Wavelength")

#Remove the first two rows, which just lists units and equations
spectra <- spectra[-c(1:2),]

#Make wavelength row name
rownames(spectra) <- spectra[,1]

#And drop the Wavelength column
spectra <- spectra[,-which(names(spectra) == "Wavelength")]

#Now convert all columns to numeric
#Spectra is still a list
#May be able to swap 1:length(filenames) with filenames; this specifies which columns to apply it to
spectra[,1:length(filenames)] <- lapply(spectra[,1:length(filenames)], as.numeric)

#Swap the columns and rows and make it a dataframe
spectra <- data.frame(t(spectra))

#Add the file names as the first column
spectra <- cbind(filenames, spectra)


#clear up the working directory by removing everything except the final data frame
rm(rawabs, filenames, colnames, i)



#Do the calculations####

#Complete the normalization to cell size
#The equation is a = 2.303A/l (what we used in the SERC paper; accepted in Helms et al 2008/Hu et al 2002)
#This is the Napierian form - Hu et al recommends reporting everything as this
#However the decadic form is used for SUVA calculations - which is raw absorbance/path length
#We will report everything in Napierian, except one a254 value in decadic
#Set the equation up as a function
#The l will be 0.01 m since we use a 1 cm cuvette
#After this, the unit will be m^-1
abs.norm <- function(x){
  (2.303*x)/0.01
}
#Now apply the function to every cell we want
spectra[,2:602] <- lapply(spectra[,2:602], abs.norm)

#set up an empty dataframe with properly named columns
columns <- c("name","decadic.a254","S275_295","S350_400","SRatio") 

# pass this vector length to ncol parameter
# and nrow with 0
calcs <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
# assign column names
colnames(calcs) = columns

#set up a dataframe for the row
row <- data.frame(matrix(nrow = 1, ncol = length(columns))) 
# assign column names
colnames(row) = columns


#Now complete the calculations for each row
for(i in unique(spectra$filenames)){
    #Pull out one spectra/observation
  obs <- filter(spectra, filenames == i)
  #Pull out decadic a254
  decadic.a254 <- (obs$X254)/2.303
  #Pull out the name
  name <- obs$filenames
  #Leave out the filenames column
  obs <- obs[,-c(1)]
  
  #calculate S275_295 and S350_400 through linear regression of log transformed spectra
  #This is the preferred method in Helms et al 2008, but this could also be done by 
  #fitting an exponential decay equation; there was less than a 1% difference between methods
  #First, use an if/else statement to put in NA for any blanks - used "lank" here so capital/lowercase B is okay
    if(grepl("lank", name)) {
      #Put in NA for all of the spectral slope calculations
      row[,1] <- name
      row[,2] <- decadic.a254
      row[,3] <- NA
      row[,4] <- NA
      row[,5] <- NA
     } else {
    
  #Then calculate everything
  #Take the log
  #NOTE: TAKING THE LOG WILL CAUSE WARNINGS FOR THE 0s/negatives, it doesn't matter
  log <- log(obs)
  #Pivot it into long format
  log <- gather(log)
  #Remove the X and convert to numeric
  log$key <- as.numeric(str_replace(log$key, "X", ""))
  
  #pull out the log 275 to 295
  l275_295 <- filter(log, key >= 275 & key <= 295)
  #linear regression
  regl275_295 <- lm(formula = l275_295$value ~ l275_295$key)
  #save the slope, multiply by -1 since the slope isn't given as a negative
  S275_295 <- -1*regl275_295$coefficients[2]
  
  #and log 350 to 400
  l350_400 <- filter(log, key >= 350 & key <= 400)
  #linear regression
  regl350_400 <- lm(formula = l350_400$value ~ l350_400$key)
  #save the slope, multiply by -1 since the slope isn't given as a negative
  S350_400 <- -1*regl350_400$coefficients[2]
  
  #Calculate the slope ratio (S275_295 to S350_400)
  SRatio <- S275_295/S350_400
  
  #Put all these values into the dataframe for all the calculations
  #Put all of these values into a row
  row[,1] <- name
  row[,2] <- decadic.a254
  row[,3] <- S275_295
  row[,4] <- S350_400
  row[,5] <- SRatio
     } #end the else
  #rbind the row onto calcs
  calcs <- rbind(calcs, row)
}

#Clean it up so you just have calcs, spectra, and what we need for EEMs section
rm(l275_295, l350_400, log, obs, regl275_295, regl350_400, row, decadic.a254, columns, i, name, S275_295, S350_400, SRatio, abs.norm)

#Bind calcs and spectra together
compilation <- cbind(calcs, spectra)
#HERE YOU CAN CHECK ALL OF THE FILE NAMES LINED UP PROPERLY

#Remove the extra column of file names

# create column for dilution factor
compilation <- select(compilation, -filenames)
compilation$dilution <- 1
for(i in 1:nrow(compilation)) {
  x <- compilation$decadic.a254[i]
  if(x > 20 && x < 30) {
    compilation$dilution[i] <- 1.6
  } else if(x > 30 && x < 40) {
    compilation$dilution[i] <- 2
  } else if(x > 40 && x < 50) {
    compilation$dilution[i] <- 2.6666666666666
  } else if(x > 50 && x < 80) {
    compilation$dilution[i] <- 4
  } else if(x > 80) {
    compilation$dilution[i] <- NA
  } else {
    compilation$dilution[i] <- 1
  }
}

#CHANGE WORKING DIRECTORY TO WHEREVER YOU WANT TO SAVE THE FILE
#Save File
#UPDATE FILE NAME AS NEEDED
write.csv(compilation, file.path(fp.results, "AbsorbanceCompilation.csv"))

#Now make a dilution factor dataframe
dilfac <- data.frame(name = compilation$name, dilution = compilation$dilution)
#And remove the A
dilfac$name <- substring(dilfac$name, 2)
```


#Fluorescence indices formatting/calculations- Calculating everything the same way as the old fluorometer
HIX, BIX, FI, P/H, Protein-like, and humic-like


```{r EEMold}
knitr::opts_chunk$set(echo = TRUE)

#Make a list of all of the filenames in the folder
filenames <- list.files(fp.eems.in, pattern="*.dat", full.names = TRUE)
#Load in all of the files to one list
allEEM <- sapply(filenames,FUN = read.delim,simplify = FALSE,USE.NAMES = TRUE)
#Need to clean up the names in the list
#Change filenames to remove the canned part we don't want
filenames <- sub(" .*", "", basename(filenames))
#Then sub the name onto the list
names(allEEM) <- filenames


#Need to load in the reference blank
blank <- read.delim(file.path(fp.data, "EBlank1 (01) - Waterfall Plot Blank.dat"))

#Calculate the parameter from the blank that we need in Protein-like/Humic-like calculations
blankem376_424 <- filter(blank, Wavelength >= 376 & Wavelength <= 424)
blanksum <- sum(blankem376_424$X350)

#For this code, at this point we will want to do the calculations for each csv and then compile it into one dataframe
#Now we'll set up a loop to do the calculations on each EEM individually and generate one dataframe of results
#We're going to need to be careful here keeping track of names, since names don't populate in the list

#set up an empty dataframes with properly named columns
columns <- c("Sample","oldFI","oldBIX","oldHIX","P/H", "Protein-like", "Humic-like") 
# pass this vector length to ncol parameter
# and nrow with 0
oldindices <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
# assign column names
colnames(oldindices) = columns

#set up a dataframe for the row
row <- data.frame(matrix(nrow = 1, ncol = length(columns))) 
# assign column names
colnames(row) = columns



for(i in unique(filenames)){
 
  #Pull out a single EEM
  oneEEM <- allEEM[[i]]
  
  #Transform the EEM, clean up the data
  #we need to remove the "nm" row 
  oneEEM <- oneEEM[-c(1),]
  #make Column 1 numeric too
  oneEEM$Sample...Blank <- as.numeric(oneEEM$Sample...Blank)
  
  #Correct for dilution, multiply EEM by dilution factor
  #Pull out the dilution factor for that sample
  onedilfac <- filter(dilfac, name == substring(i, 2))
  onedilfac <- onedilfac$dilution
  
  #And multiply the EEM by it, excluding the column with the wavelength
  oneEEM <- oneEEM %>% mutate(across(starts_with("X"), ~ . *onedilfac))
  
  #FI
  #Originally in McKnight et al 2001 excitation at 370 with ratio between 450nm:500nm (included what it is in our old python code)
  #Note, in the new EEMS this should be updated as Cory et al 2010
  em450 <- filter(oneEEM, Sample...Blank == 448)
  em500 <- filter(oneEEM, Sample...Blank == 498)
  oldFI <- em450$X370/em500$X370
  rm(em450, em500)
  
  #BIX
  #For excitation at 310 nm, ratio of emission 380nm/430nm (USGS protocol)
  em380 <- filter(oneEEM, Sample...Blank == 380)
  em430 <- filter(oneEEM, Sample...Blank == 430)
  oldBIX <- em380$X310/em430$X310
  rm(em380, em430)
  
  #Now calculate HIX, this is based on Ohno, 2002; Zsolnay et al., 1999; and listed in USGS Aqualog Protocol
  #Excitation wavelength of 254 -> we just need this column, will use 255
  #sum 435 to 480, sum 300 to 345
  em435_480 <- filter(oneEEM, Sample...Blank >= 435 & Sample...Blank <= 480)
  em300_345 <- filter(oneEEM, Sample...Blank >= 300 & Sample...Blank <= 345)
  oldHIX <- sum(em435_480$X255)/(sum(em435_480$X255)+sum(em300_345$X255))
  rm(em435_480,em300_345)
  
  
  #P/H
  #ratio between 340 nm emission @ 275 nm excitation: 480 nm emission @ 350 nm excitation (old python) 
  #NEED TO FIND SOURCE
  em340 <- filter(oneEEM, Sample...Blank == 340)
  em480 <- filter(oneEEM, Sample...Blank == 480)
  PH <- em340$X275/em480$X350

  
  #Protein-like
  #Source: Old python code
  #(ex274/em340)/(sum of emission wavelengths between 376 nm to 424 nm for excitation 350 nm for the BLANK) (old python) 
  pl <- em340$X275/blanksum
  
  
  #Humic-like
  #Source: old python code
  #(ex350/em480)/(sum of emission wavelengths between 376 nm to 424 nm for excitation 350 nm for the BLANK) (old python) 
  hl <- em480$X350/blanksum
  
  
  rm(em340, em480)
  
  #Put all these values into the dataframe for all the calculations
  #Put all of these values into the row
  row$Sample <- i
  row$oldFI <- oldFI
  row$oldBIX <- oldBIX
  row$oldHIX <- oldHIX
  row$`P/H`<-PH
  row$`Protein-like`<-pl
  row$`Humic-like`<-hl
  
  #rbind the row onto calcs
  oldindices <- rbind(oldindices, row)
   
}

#Clean up the environment
rm(oneEEM, row, oldBIX, columns, oldFI, oldHIX, filenames, hl, i, PH, pl, onedilfac)


#CHANGE WORKING DIRECTORY TO WHEREVER YOU WANT TO SAVE THE FILE
#Save File
#UPDATE FILE NAME AS NEEDED
write.csv(oldindices, file.path(fp.results,"FluoroOldIndices.csv"))

#clean up the file names
rm(allEEM, blank, blankem376_424, blanksum, x)

```


#Corrections and indices calculations with staRdom

Inner filter effect correction, raman normalization, remove raman/rayleigh scattering lines, and calculate indices using staRdom
Calculate all indices in staRdom (FI, HIX, BIX, and Coble's peaks)

```{r EEMnew}

#Read in EEMs - this will read in every file that's in that folder
#Read in all the files that are in your EEM folder
#It works for aqualog files
eem_list <- eem_read(fp.eems.raw, recursive = TRUE, import_function = eem_csv) 

#Read in absorbance in low resolution absorbance folder - these are the ones used for the inner filter effect
#Then, will need to format to the same format as staRdom requires
#Call in filenames
filenames <- list.files(fp.lowa, pattern="*.dat", full.names = TRUE)
# then call in spectra
rawabs <- sapply(filenames, FUN = read.delim, simplify = FALSE, USE.NAMES = TRUE)
# Need to clean up the names in the list
# Change filenames to remove the canned part we don't want
# then make these object names in rawabs
filenames <- sub(" .*", "", basename(filenames))
names(rawabs) <- filenames
# Now, compile all of these columns into one dataframe
# Remove all columns except for the absorbance
rawabs <- lapply(rawabs, function(x) {select(x, -c("I1", "I1.dark", "R1", "R1dark", "XCorrect", "I1c", "R1c", "I1c.R1c", "Percent.T"))})
#Make the Absorbance column name the sample name and remove extra columns
for(i in unique(1:length(filenames))){
  colnames <- c("Wavelength", filenames[i])
  colnames(rawabs[[i]]) <- colnames
}
#Combine into one dataframe, joining by Wavelength column
absorbance <- reduce(rawabs, left_join, by = "Wavelength")
#Remove the first two rows, which just lists units and equations
absorbance <- absorbance[-c(1:2),]
#And make absorbance numeric
absorbance[,] <- lapply(absorbance[,], as.numeric)
#Change name of Wavelength so "W" is lowercase
absorbance <- absorbance %>% rename_at('Wavelength', ~'wavelength')
#Arrange so it's listed by wavelength from low to high
absorbance <- arrange(absorbance, wavelength)


#Trying to format absorbance data to be read in by staRdom
#Read in absorbance in low resolution absorbance folder - these are the ones used for the inner filter effect
#Need to convert to low resolution absorbance data to a format staRdom can understand
#We need a csv for each sample where there's one wavelength column and one with the absorbance
#Then, read in with staRdom
#Call in filenames
filenames <- list.files(fp.lowa, pattern="*.dat", full.names = TRUE)
# then call in spectra



#Make our metatable
metatable <- system.file("extdata/metatable_dreem.csv",package = "staRdom") # path to example data, can be replaced by a path to your own data
metaexample <- read.table(metatable, header = TRUE, sep = ",", dec = ".", row.names = 1) # load data

#Or make a metatable
eem_metatemplate(eem_list, absorbance) %>%
  write.csv(file="metatable.csv", row.names = FALSE)
#We will make something like this
#Since the EEMs are a list, I think we can get everything to this point on our own
#May need to pivot longer the EEMs
problem <- eem_checkdata(eem_list,absorbance,meta,metacolumns = c("dilution"),error=FALSE)

#Fix sample names
eem_list <- eem_name_replace(eem_list,c("\\(FD3\\)"),c(""))

#IFE correction
eem_list <- eem_ife_correction(eem_list,absorbance, cuvl = 5)
#Need absorbance to be in a dataframe where the first column is the wavelength from low to high, and each other column is the sample named with sample name
#Need EEM in a list: sample = character of 1 with sample name; ex = double with excitation wavelengths; em = integer with emission wavelengths;
#x = intensity (ours actually may not look that different)
#Our cuvl = 1, 1cm cuvette

#plot the EEMs
eem_overview_plot(eem_list, spp=9, contour = TRUE)

#Raman normalization:
eem_list <- eem_raman_normalisation2(eem_list, blank = "blank")

#Remove blanks from EEM and absorbance sample lists
eem_list <- eem_extract(eem_list, c("nano", "miliq", "milliq", "mq", "blank"),ignore_case = TRUE)
absorbance <- dplyr::select(absorbance, -matches("nano|miliq|milliq|mq|blank", ignore.case = TRUE))

#Remove and interpolate scattering
remove_scatter <- c(TRUE, TRUE, TRUE, TRUE)
remove_scatter_width <- c(15,15,15,15)

eem_list <- eem_rem_scat(eem_list, remove_scatter = remove_scatter, remove_scatter_width = remove_scatter_width)

#Interpolate EEMs across scattering lines
eem_list <- eem_interp(eem_list, cores = cores, type = 1, extend = FALSE)

#Correction for dilution
dil_data <- meta["dilution"]
#Row name is sample name, the value in the one column is the dilution factor, could pull this from metadata sheet
eem_list <- eem_dilution(eem_list,dil_data)

#Smooth data for peak picking
eem4peaks <- eem_smooth(eem_list, n = 4, cores = cores)

#Summarize the samples and the corrections that have been done
summary(eem_list)

#And pull out all of the indices and peaks
bix <- eem_biological_index(eem4peaks)
coble_peaks <- eem_coble_peaks(eem4peaks)
fi <- eem_fluorescence_index(eem4peaks)
hix <- eem_humification_index(eem4peaks, scale = TRUE)

indices_peaks <- bix %>%
  full_join(coble_peaks, by = "sample") %>%
  full_join(fi, by = "sample") %>%
  full_join(hix, by = "sample")

indices_peaks

```